---
title: "Limited Dependent Varible Models Page"
author: "Kinito Solutions"
date: "September 11, 2016"
output: 
  html_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE);
library(VGAM);
library(AER);
library(truncreg);
library(censReg);
```

# Limited Dependent Varible Models

A limited dependent variable means that there is a limit or boundary on the dependent variable and some of the observations “hit” this limit.   A limited dependent variable is a continuous variable with a lot of repeated observations at the lower or upper limit.  Examples include the quantity of a product consumed, the number of hours that women work, etc. 

Limited dependent variable models address two issues: censoring and truncation.  Censoring is when the limit observations are in the sample and truncation is when the observations are not in the sample. 

# Limited Dependent Varible Models Example

```{r Limited Dependent Varible Models example}
myDataFrame<- read.csv("Data/limdep_ambexp.csv")
head(myDataFrame)
attach(myDataFrame)
# Define variables
Y <- cbind(ambexp)
X <- cbind(age, female, totchr)
# Descriptive statistics
summary(Y)
summary(X)
# Tobit model coefficients (sensoring from below at 0)
myTobitModel <- tobit(Y ~ X, left=0, right=Inf, data=myDataFrame)
summary(myTobitModel)
myTobitCensored <- censReg(Y ~ X, left=0, right=Inf, data=myDataFrame)
summary(myTobitCensored)
# Tobit model marginal effects for the censored sample
summary(margEff(myTobitCensored))
# Probit model coefficients
myProbitModel <- glm(I(Y > 0) ~ X, data = myDataFrame, family = binomial(link = "probit"))
summary(myProbitModel)
# Truncated regression coefficients
myDataFrameSubset <- subset(myDataFrame, Y>0)
#myLeftTruncatedRegression <- truncreg(Y ~ X, point = 0, direction = "left", data=myDataFrameSubset)
#summary(myLeftTruncatedRegression)
# Test for Tobit versus Probit and Truncated regression
#lrtest <- 2*(logLik(myProbitModel) + logLik(myLeftTruncatedRegression) - logLik(myTobitModel))
#lrtest
```
