---
title: "Multinomial Probit and Logit Models Page"
author: "Kinito Solutions"
date: "September 11, 2016"
output: 
  html_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE);
library(mlogit);
```

# Multinomial Probit and Logit Models

An additional way to get slope coefficients: Dividing the sum of the residuals of the regression of x1 on all the other independent variables multiplied by the outcome variable (dep. variable) throught the sum of those squared residuals. To express and check this in R should pose no problem since we already know commands that are needed for that:

# Multinomial Probit and Logit Models Example

```{r Multinomial Probit and Logit Models example}
myDataFrame<- read.csv("Data/multinomial_fishing1.csv")
head(myDataFrame)
attach(myDataFrame)
# Descriptive statistics
table(mode)
# Reshaping the data from wide to long format
#myDataFrame$mode<-as.factor(myDataFrame$mode)
mlogitData <- mlogit.data(myDataFrame,
                          varying=4:15,
                          choice="mode",
                          shape="wide")
mlogitData[1:20,]
# Multinomial logit model coefficients 
mlogitModelCharter <- mlogit(mode ~ 1 | income,
              							data = mlogitData,
              							reflevel="charter")
summary(mlogitModelCharter)
# Multinomial logit model coefficients (with different base outcome)
mlogitModelPier <- mlogit(mode ~ 1 | income,
            							data = mlogitData,
            							reflevel="pier")
summary(mlogitModelPier)
# Multinomial logit model odds ratios 
exp(coef(mlogitModelCharter))
# Conditional logit model
mlogitConditionalModelCharter <- mlogit(mode ~ price+catch | income,
                          							data = mlogitData,
                          							reflevel="charter")
summary(mlogitConditionalModelCharter)
mlogitConditionalModelPier <- mlogit(mode ~ price+catch | income,
                  									data = mlogitData,
                  									reflevel="pier")
summary(mlogitConditionalModelPier)
# Setting mean values for variables to use for marginal effects 
m <- mlogit(mode ~ price+catch | income,
      			data = mlogitData,
      			reflevel="charter")
z <- with(mlogitData,
			data.frame(price = tapply(price, index(m)$alt, mean), catch = tapply(catch, index(m)$alt, mean), income = mean(income))
		)
# Multinomial logit model marginal effects
effects(mlogitModelCharter, covariate = "income", data = z)
# Conditional logit model marginal effects
effects(mlogitConditionalModelCharter, covariate = "income", data = z)
effects(mlogitConditionalModelCharter, covariate = "price", data = z)
effects(mlogitConditionalModelCharter, covariate = "catch", data = z)
# Multinomial probit model coefficients 
#mprobit.model1 <- mlogit(mode ~ 1 | income, data = mlogitData, reflevel="charter", probit=TRUE)
#summary(mprobit.model1)
# Hauseman-McFadden test of independence of irrelevant alternatives
mSet <- mlogit(mode ~ 1 | income,
               data = mlogitData,
               reflevel="beach")
mSubset <- mlogit(mode ~ 1 | income,
                  data = mlogitData,
                  reflevel="beach",
                  alt.subset=c("beach", "pier", "private"))
hmftest(mSet, mSubset)
```
