---
title: "Panel Data Models Page"
author: "Kinito Solutions"
date: "September 11, 2016"
output: 
  html_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE);
library("foreign");
library(plm);
```

# Panel Data Models

Panel data models provide information on individual behavior, both across individuals and over time.  The data and models have both cross-sectional and time-series dimensions.  Panel data can be balanced when all individuals are observed in all time periods or unbalanced when individuals are not observed in all time periods. Examples include estimating the effect of education on income, with data across time and individuals; and estimating the effects of income on savings, with data across years and countries.

# Panel Data Models Example

```{r Panel Data Models example}
myDataFrame <- read.csv("Data/panel_wage.csv")
head(myDataFrame);
attach(myDataFrame)
Y <- cbind(lwage)
X <- cbind(exp, exp2, wks, ed)
# Set data as panel data
myPanelData <- plm.data(myDataFrame, index=c("id","t"))
# Descriptive statistics
summary(Y)
summary(X)
# Pooled OLS estimator
myPoolingOLSEstimator <- plm(Y ~ X, data=myPanelData, model= "pooling")
summary(myPoolingOLSEstimator)
# Between estimator
myBetweenEstimator <- plm(Y ~ X, data=myPanelData, model= "between")
summary(myBetweenEstimator)
# First differences estimator
myFirstDifferencesEstimator <- plm(Y ~ X, data=myPanelData, model= "fd")
summary(myFirstDifferencesEstimator)
# Fixed effects or within estimator
myWithinEstimator <- plm(Y ~ X, data=myPanelData, model= "within")
summary(myWithinEstimator)
# Random effects estimator
myRandomEstimator <- plm(Y ~ X, data=myPanelData, model= "random")
summary(myRandomEstimator)
# LM test for random effects versus OLS
plmtest(myPoolingOLSEstimator)
# LM test for fixed effects versus OLS
pFtest(myWithinEstimator, myPoolingOLSEstimator)
# Hausman test for fixed versus random effects model
phtest(myRandomEstimator, myWithinEstimator)
```
