---
title: "Probit and Logit Models Page"
author: "Kinito Solutions"
date: "September 11, 2016"
output: 
  html_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE);
```

# Probit and Logit Models

Probit and logit models are among the most popular models.  The dependent variable is a binary response, commonly coded as a 0 or 1 variable.  The decision/choice is whether or not to have, do, use, or adopt.  Examples include whether a consumer makes a purchase or not, and whether an individual participates in the labor market or not.

# Probit and Logit Models Example

```{r Probit and Logit Models example}
myDataFrame<- read.csv("Data/probit_insurance.csv")
head(myDataFrame)
attach(myDataFrame)
# Define variables
Y <- cbind(ins)
X <- cbind(retire, age, hstatusg, hhincome, educyear, married, hisp)
# Descriptive statistics
summary(Y)
summary(X)
table(Y)
table(Y)/sum(table(Y))
# Regression coefficients
olsRegressionModel <- lm(Y ~ X)
summary(olsRegressionModel)
# Logit model coefficients
logitModelCoefs <- glm(Y ~ X, family=binomial (link = "logit"))
summary(logitModelCoefs) 
# Logit model odds ratios
exp(logitModelCoefs$coefficients)
# Probit model coefficients
probitModelCoefs <- glm(Y ~ X, family=binomial (link="probit"))
summary(probitModelCoefs)
# Regression marginal effects
coef(olsRegressionModel)
# Logit model average marginal effects
LogitScalar <- mean(dlogis(predict(logitModelCoefs, type = "link")))
LogitScalar * coef(logitModelCoefs)
# Probit model average marginal effects
ProbitScalar <- mean(dnorm(predict(probitModelCoefs, type = "link")))
ProbitScalar * coef(probitModelCoefs)
# Regression predicted probabilities
olsRegressionProbs <- predict(olsRegressionModel)
summary(olsRegressionProbs)
# Logit model predicted probabilities
logitModelProbs <- predict(logitModelCoefs, type="response")
summary(logitModelProbs)
# Probit model predicted probabilities
probitModelProbs <- predict(probitModelCoefs, type="response")
summary(probitModelProbs)
# Percent correctly predicted values
table(true = Y, pred = round(fitted(probitModelCoefs)))
table(true = Y, pred = round(fitted(logitModelCoefs))) 
# McFadden's Pseudo R-squared
probitModelCoefsUpdate <-update(probitModelCoefs, formula= Y ~ 1)
McFadden<- 1-as.vector(logLik(probitModelCoefs)/logLik(probitModelCoefsUpdate))
McFadden
```
