---
title: "Survival Analysis Page"
author: "Kinito Solutions"
date: "September 11, 2016"
output: 
  html_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE);
library(survival);
```

# Survival Analysis

Survival analysis is applied when the data set includes subjects that are tracked until an event happens (failure) or we lose them from the sample.  We are interested in how long they stay in the sample (survival).  We are also interested in their risk of failure (hazard rates).  Examples include loan performance and default, firm survival and exit, and time to retirement.

# Survival Analysis Example

```{r Survival Analysis example}
myDataFrame<- read.csv("Data/survival_unemployment.csv")
head(myDataFrame)
attach(myDataFrame)
# Define variables 
timeList <- spell
eventList <- event
X <- cbind(logwage, ui, age)
groupList <- ui
# Descriptive statistics
summary(timeList)
summary(eventList)
summary(X)
summary(groupList)
# Kaplan-Meier non-parametric analysis
nonParamKMSurvival <- survfit(Surv(timeList,eventList) ~ 1)
summary(nonParamKMSurvival)
plot(nonParamKMSurvival, xlab="Time", ylab="Survival Probability")
# Kaplan-Meier non-parametric analysis by Group
nonParamKMSurvivalByGroup <- survfit(Surv(timeList, eventList) ~ groupList)
summary(nonParamKMSurvivalByGroup)
plot(nonParamKMSurvivalByGroup, xlab="Time", ylab="Survival Probability")
# Nelson-Aalen non-parametric analysis
nonParamNASurvival <- survfit(coxph(Surv(timeList,eventList)~1), type="aalen")
summary(nonParamNASurvival)
plot(nonParamNASurvival, xlab="Time", ylab="Survival Probability")
# Cox proportional hazard model - coefficients and hazard rates
coxProportionalHazardBreslow <- coxph(Surv(timeList,eventList) ~ X, method="breslow")
summary(coxProportionalHazardBreslow)
# Exponential, Weibull, and log-logistic parametric model coefficients
# Opposite signs from Stata results, Weibull results differ; same as SAS
exponentialSurvival <- survreg(Surv(timeList,eventList) ~ X, dist="exponential")
summary(exponentialSurvival)
weibullSurvival <- survreg(Surv(timeList,eventList) ~ X, dist="weibull")
summary(weibullSurvival)
loglogisticSurvival <- survreg(Surv(timeList,eventList) ~ X, dist="loglogistic")
summary(loglogisticSurvival)
```
