---
title: "Time Series ARIMA Models Page"
author: "Kinito Solutions"
date: "September 11, 2016"
output: 
  html_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE);
library(timeSeries);
```

# Time Series ARIMA Models

Time series ARIMA models are applied with time series data of variables measured over time.  Time series analysis examines relationships of variables over time such as commodity prices or crop yields.  Time series models may be used for analyzing the effects of a specific event (such as the effects of the recession on unemployment rates) or for forecasting (for example to predict economic growth or future prices).

# Time Series ARIMA Models Example

```{r Time Series ARIMA Models example}
myDataFrame<- read.csv("Data/timeseries_ppi.csv")
head(myDataFrame)
attach(myDataFrame)
# Defining variables
Y <- ppi
diffY <- diff(Y)
t <- yearqrt
# Descriptive statistics and plotting the data
summary(Y)
summary(diffY)
plot(t,Y)
plot(diffY)
# Dickey-Fuller test for variable
#adf.test(Y, alternative="stationary", k=0)
#adf.test(Y, alternative="explosive", k=0)
#summary(lm(dppi ~ lppi, na.action=na.omit))
#summary(lm(dppi ~ lppi + trend, na.action=na.omit))
# Augmented Dickey-Fuller test
#adf.test(Y, alternative="stationary")
# DF and ADF tests for differenced variable
#adf.test(d.Y, k=0)
#adf.test(d.Y)
# ACF and PACF
acf(Y)
pacf(Y)
acf(diffY)
pacf(diffY)
# ARIMA(1,0,0) or AR(1)
arima(Y, order = c(1,0,0))
# ARIMA(2,0,0) or AR(2)
arima(Y, order = c(2,0,0))
# ARIMA(0,0,1) or MA(1)
arima(Y, order = c(0,0,1))
# ARIMA(1,0,1) or AR(1) MA(1)
arima(Y, order = c(1,0,1))
# ARIMA on differenced variable 
# ARIMA(1,1,0)
arima(diffY, order = c(1,0,0))
# ARIMA(0,1,1)
arima(diffY, order = c(0,0,1))
# ARIMA(1,1,1)
arima(diffY, order = c(1,0,1))
# ARIMA(1,1,3)
arima(diffY, order = c(1,0,3))
# ARIMA(2,1,3)
arima(diffY, order = c(2,0,3))
# ARIMA(1,0,1) forecasting
myARIMA101 <- arima(Y, order = c(1,0,1))
myDataFrame.pred1 <- predict(myARIMA101, n.ahead=100)
plot(Y)
lines(myDataFrame.pred1$pred, col="blue")
lines(myDataFrame.pred1$pred+2*myDataFrame.pred1$se, col="red")
lines(myDataFrame.pred1$pred-2*myDataFrame.pred1$se, col="orange")
# ARIMA(1,1,1) forecasting
myARIMA111 <- arima(diffY, order = c(1,1,1))
myDataFrame.pred1 <- predict(myARIMA111, n.ahead=100)
plot(diffY)
lines(myDataFrame.pred1$pred, col="blue")
lines(myDataFrame.pred1$pred+2*myDataFrame.pred1$se, col="red")
lines(myDataFrame.pred1$pred-2*myDataFrame.pred1$se, col="orange")
```
